### 2021-10-07 proof/document/root.tex
Move proof files to subdirectory

\documentclass[10pt]{article}
\usepackage{isabelle,isabellesym}
\usepackage[a4paper,portrait,margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{pdfsetup}

\urlstyle{rm}
\isabellestyle{tt}

\begin{document}
\title{A highly-available move operation for replicated trees (Proof Document)}
\author{Martin Kleppmann, Dominic P. Mulligan, Victor B. F. Gomes, and Alastair R. Beresford}
\date{}
\maketitle
\tableofcontents
\parindent 0pt\parskip 0.5ex

\input{session} % generated text of all theories
\end{document}